---
output: github_document
---

# S-MiXcan

**S-MiXcan** is an R package for performing cell-type-aware **summary-based transcriptome-wide association studies (TWAS)**. It extends the MiXcan framework to analyze associations between genetically regulated gene expression (GReX) and traits and their functional cell types using GWAS summary statistics.

------------------------------------------------------------------------

## 🔬 Overview

Traditional TWAS approaches predict gene expression at the *bulk tissue* level and test its association with disease traits, ignoring heterogeneity across cell types. **MiXcan** improves on this by enabling cell-type-aware TWAS, but it requires *individual-level genotype data*.

**S-MiXcan** addresses this limitation.\
It provides a **summary-statistics-based** framework that:

-   Infers cell-type-specific GReX--trait associations from GWAS summary statistics.
-   Leverages MiXcan-trained cell-type models.
-   Evaluates functional cell-types.
-   Requires **no individual-level genotype data**.

------------------------------------------------------------------------

## 📦 Installation


You can install the development version of `S-MiXcan` from GitHub:

```{r, eval = TRUE}
# Install devtools if not already installed
if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}

# Install S-MiXcan from GitHub
devtools::install_github("songxiaoyu/SMiXcanK")

library(SMiXcanK)
```

------------------------------------------------------------------------

## 🧪 Example Usage

```{r, eval = TRUE}
# Define input 
data(W)      
str(W)
data(gwas_results)    
data(X_ref_filtered)   
data(n0)
data(n1)


# Perform association test
S_MiXcan_results <- SMiXcan_assoc_test(W[,1], W[,2], gwas_results, X_ref_filtered, n0=n0, n1=n1, family='binomial')
```

------------------------------------------------------------------------

## 📄 Citation

If you use **S-MiXcan** in your research, please cite this page

------------------------------------------------------------------------

## 📫 Contact

For questions, please contact:

**Sinan Zhu**\
PhD Candidate, Duke-NUS Medical School\
Email: [sinan.zhu\@u.duke.nus.edu](mailto:sinan.zhu@u.duke.nus.edu){.email}

------------------------------------------------------------------------

## 🔒 License

------------------------------------------------------------------------

\`\`\`r rmarkdown::render("README.Rmd")
