---
output: github_document
---

# S-MiXcan

**S-MiXcan** is an R package for performing cell-type-aware **summary-based transcriptome-wide association studies (TWAS)**. It extends the MiXcan framework to analyze associations between genetically regulated gene expression (GReX) and traits and their functional cell types using GWAS summary statistics.

------------------------------------------------------------------------

## 🔬 Overview

Traditional TWAS approaches predict gene expression at the *bulk tissue* level and test its association with disease traits, ignoring heterogeneity across cell types. **MiXcan** improves on this by enabling cell-type-aware TWAS, but it requires *individual-level genotype data*.

**S-MiXcan** addresses this limitation.\
It provides a **summary-statistics-based** framework that:

-   Infers cell-type-specific GReX--trait associations from GWAS summary statistics.
-   Leverages MiXcan-trained cell-type models.
-   Evaluates functional cell-types.
-   Requires **no individual-level genotype data**.

------------------------------------------------------------------------

## 📦 Installation

You can install the development version of `S-MiXcan` from GitHub:

```{r, eval = FALSE}
# Install devtools if not already installed
if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}

# Install S-MiXcan from GitHub
devtools::install_github("songxiaoyu/SMiXcanK")

library(SMiXcanK)
```

------------------------------------------------------------------------

## 🧪 Example Usage

```{r, eval = TRUE}
# Define input 
## Example: regularized_inverse_cov()

set.seed(1)

# Create a well-behaved covariance matrix (p x p)
p <- 5
A <- matrix(rnorm(p * p), p, p)
X <- crossprod(A)  # symmetric positive-definite covariance-like matrix

# Regularized inverse
ri <- regularized_inverse_cov(X)

# Outputs
ri$lambda        # regularization strength used (scalar here)
ri$inv[1:3, 1:3] # top-left corner of the inverse matrix


## Example: SMiXcan_assoc_test_K() with K = 3 cell types

set.seed(123)

# Dimensions
n  <- 400   # reference panel individuals
p  <- 60    # SNPs in gene region
K  <- 3     # cell types

# Reference genotype matrix (n x p); here simulated as standardized normals
x_g <- scale(matrix(rnorm(n * p), n, p))

# Cell-type weights matrix (p x K)
W <- matrix(rnorm(p * K), p, K)

# GWAS summary stats for the same p SNPs
gwas_results <- list(
  Beta    = rnorm(p, mean = 0, sd = 0.05),
  se_Beta = runif(p, min = 0.02, max = 0.08)
)

# Case/control sample sizes (for binomial null model)
n1 <- 5000
n0 <- 5000

# Run S-MiXcan association test (K = 3)
res <- SMiXcan_assoc_test_K(
  W = W,
  gwas_results = gwas_results,
  x_g = x_g,
  n0 = n0,
  n1 = n1,
  family = "binomial"
)

# Results:
# - Z_join: joint-model Z for each cell type (length K)
# - p_join_vec: per-cell-type joint p-values (length K)
# - p_join: ACAT-combined p across K cell types
res$Z_join
res$p_join_vec
res$p_join

```

------------------------------------------------------------------------

## 📄 Citation

If you use **S-MiXcan** in your research, please cite this page

------------------------------------------------------------------------

## 📫 Contact

For questions, please contact:

**Sinan Zhu**\
PhD Candidate, Duke-NUS Medical School\
Email: [sinan.zhu\@u.duke.nus.edu](mailto:sinan.zhu@u.duke.nus.edu){.email}

------------------------------------------------------------------------

## 🔒 License

------------------------------------------------------------------------

\`\`\`r rmarkdown::render("README.Rmd")
